Creating the ``r params$package_name`` R package

### 2025-10-08 create-FSHybridPLS.Rmd
scalar normalization test done

The class definition is a simple list that contains a matrix and a collection of \code{fd} objects from the \code{fda} package, along with the necessary metadata.
- gram matrices are saved as a list.
```{r classdef:predictor_hybrid}

#' Create a predictor_hybrid object (S3 version, automatic basis size and Gram matrix)
#'
#' Constructs a hybrid predictor object that stores both scalar and functional predictors,
#' and automatically computes Gram matrices (inner products of basis functions).
#'
#' @param Z A numeric matrix of dimension \code{n_sample × n_scalar} representing scalar predictors.
#' @param functional_list A list of functional predictors (e.g., \code{fd} objects from the \code{fda} package).
#'
#' @return An object of class \code{predictor_hybrid}, containing scalar and functional data with Gram matrices.
#' @export
predictor_hybrid <- function(Z, functional_list) {
  stopifnot(is.matrix(Z), is.numeric(Z))
  stopifnot(is.list(functional_list))

@@

```{r}
#' Construct a Single-Sample Predictor Hybrid Object from Coefficients
#'
#' Reconstructs a \code{predictor_hybrid} object representing one observation, using a numeric
#' coefficient vector. This alternative constructor maps the coefficients back into their functional
#' and scalar predictor representations based on the structure of a template \code{predictor_hybrid} object.
#'
#' @param format A \code{predictor_hybrid} object that provides the structure and basis information.
#' @param coef A numeric vector containing coefficients for both functional and scalar predictors.
#'
#' @return A \code{predictor_hybrid} object with updated \code{functional_list}, \code{Z}, and \code{n_sample = 1}.
#'
#'
#' @export
predictor_hybrid_from_coef <- function(format, coef){
  M <- format$n_basis_list[1] # number of basis functions
  K <- format$n_functional # number of functional predictors
  for (ii in 1:K){
    format$functional_list[[ii]] <- fd(
      coef = as.matrix(coef[((ii-1)*M + 1):(ii*M)]),
      basisobj = format$functional_list[[ii]]$basis
    )
  }
  format$Z <- t(as.matrix(coef[(K * M + 1):length(coef)]))
  format$n_sample <- 1
  return(format)
}
```

@@

@@

### curve_normalize
Subtracts the mean function and divides by a scaling constant (deno). This implementation uses the native fda S3 methods (minus.fd and times.fd).
```{r}
#' Normalize a functional data object by its mean and standard deviation.
#'
#' @param functional fd object (fda package). The functional data to be normalized (N samples).
#' @param mean_functional fd object (fda package). The mean function (1 sample).
#' @param deno Numeric scalar. The standard deviation or scaling factor.
#' @return A new fd object containing the normalized functions.
#' @export
curve_normalize <- function(functional, mean_functional, deno) {
  # Type checks
  stopifnot(
    "'functional' must be an 'fd' object" = inherits(functional, "fd"),
    "'mean_functional' must be an 'fd' object" = inherits(mean_functional, "fd"),
    "'deno' must be a single positive numeric value" = is.numeric(deno) && length(deno) == 1 && deno > 0
  )

  # 1. Centering: functional - mean_functional
  # fda::minus.fd handles the broadcasting of the single-sample mean_functional.
  centered_functional <- fda::minus.fd(functional, mean_functional)

  # 2. Scaling: (1/deno) * centered_functional
  functional_normalized <- fda::times.fd(1/deno, centered_functional)

  return(functional_normalized)
}
```

```{r}
#' 
#' @export
curve_normalize <- function(functional, mean_functional, deno){
  # operation also takes care of the original_X

@@

```{r}
#' Normalize Functional Predictors in Hybrid Objects (Train/Test Split)
#'
#' Normalizes the functional components of both the training and testing 
#' `predictor_hybrid` objects using the mean and standard deviation derived
#' ONLY from the training data. The scalar components (Z) are left unchanged.
#'
#' @param train A `predictor_hybrid` object representing the training set.
#' @param test A `predictor_hybrid` object representing the testing set.
#' @return A list containing:
#'   \item{train}{The normalized training `predictor_hybrid` object.}
#'   \item{test}{The normalized testing `predictor_hybrid` object.}
#'   \item{deno_train}{A numeric vector of scaling factors used for each functional predictor.}
#' @export
curve_normalize_train_test <- function(train, test) {
  # --- Type Checks ---
  stopifnot(
    "'train' must be of class 'predictor_hybrid'" = inherits(train, "predictor_hybrid"),
    "'test' must be of class 'predictor_hybrid'" = inherits(test, "predictor_hybrid"),
    "Mismatch in number of functional predictors" = train$n_functional == test$n_functional
  )

@@

```{r}
# scalar_normalize <- function(scalar_predictor, mean, deno){
#   scalar_predictor_normalized <- subtr_broadcast(scalar_predictor, mean) #numerator
#   deno_mat <- matrix(rep(1,nrow(scalar_predictor))) %*% deno
#   scalar_predictor_normalized <- scalar_predictor_normalized / deno_mat
#   return(scalar_predictor_normalized)
# }
#
# scalar_normalize_train_test <- function(predictor_train, predictor_test){
#   predictor_train_normalized <- predictor_train
#   predictor_test_normalized <- predictor_test
#
#   mean_train <- get_mean(predictor_train@Z) # train mean function
#   sd_train <- get_sd(predictor_train@Z)
#
#   predictor_train_normalized@Z <-
#     scalar_normalize(
#       predictor_train@Z,
#       mean_train,
#       sd_train
#       )
#
#   predictor_test_normalized@Z <-
#     scalar_normalize(
#       predictor_test@Z,
#       mean_train,
#       sd_train
#     )
#
#   return(
#     list(
#       train = predictor_train_normalized,
#       test = predictor_test_normalized
#     )
#   )
# }
```

>>>>

The class definition is a simple list that contains a matrix and a collection of \code{fd} objects from the \code{fda} package, along with the necessary metadata.
- gram matrices are saved as a list.
```{r classdef:predictor_hybrid}
#' Create a predictor_hybrid object (S3 version, automatic basis size and Gram matrix)
#'
#' Constructs a hybrid predictor object that stores both scalar and functional predictors,
#' and automatically computes Gram matrices (inner products of basis functions).
#'
#' @param Z A numeric matrix of dimension \code{n_sample × n_scalar} representing scalar predictors.
#' @param functional_list A list of functional predictors (e.g., \code{fd} objects from the \code{fda} package).
#'
#' @return An object of class \code{predictor_hybrid}, containing scalar and functional data with Gram matrices.
#' @export
predictor_hybrid <- function(Z, functional_list) {
  stopifnot(is.matrix(Z), is.numeric(Z))
  stopifnot(is.list(functional_list))

@@

```{r}
#' Construct a Single-Sample Predictor Hybrid Object from Coefficients
#' Reconstructs a \code{predictor_hybrid} object representing one observation, using a numeric
#' coefficient vector. This alternative constructor maps the coefficients back into their functional
#' and scalar predictor representations based on the structure of a template \code{predictor_hybrid} object.
#'
#' @param format A \code{predictor_hybrid} object that provides the structure and basis information.
#' @param coef A numeric vector containing coefficients for both functional and scalar predictors.
#'
#' @return A \code{predictor_hybrid} object with updated \code{functional_list}, \code{Z}, and \code{n_sample = 1}.
#'
#'
#' @export
predictor_hybrid_from_coef <- function(format, coef){
  M <- format$n_basis_list[1] # number of basis functions
  K <- format$n_functional # number of functional predictors
  for (ii in 1:K){
    format$functional_list[[ii]] <- fd(
      coef = as.matrix(coef[((ii-1)*M + 1):(ii*M)]),
      basisobj = format$functional_list[[ii]]$basis
    )
  }
  format$Z <- t(as.matrix(coef[(K * M + 1):length(coef)]))
  format$n_sample <- 1
  return(format)
}
```

@@

### add_broadcast(for matrix)

```{r}
#' @export
#' 
#' 
# --- 1. Define the S4 Generic Function ---
# This must be run first to tell R that 'add_broadcast' is a function 
# that can have methods defined for different classes.
setGeneric("add_broadcast", function(input, other, alpha = 1) {
  standardGeneric("add_broadcast")
})

# --- 2. Define the S4 Method for 'matrix' Class ---
setMethod("add_broadcast", "matrix",
###########################################
function(input, other, alpha = 1){

  # 1. Structural Check: Ensure RHS is a single observation.
  # This uses length() for vectors, and dim()[1] for matrices.
  is_matrix_other <- is.matrix(other)
  if (is_matrix_other && dim(other)[1] > 1) {
    stop("RHS must have a single observation (row).")
  }

  # 2. Conversion: Convert a single-row matrix into a vector for broadcasting.
  # If 'other' is already a vector, this does nothing.
  # If 'other' is a 1xP matrix, this flattens it into a length-P vector.
  if (is_matrix_other) {
    other <- as.vector(other) 
  }

  # 3. Native Arithmetic: Use simple R arithmetic for high performance.
  # R broadcasts the vector 'other' across the rows of the matrix 'input'.
  return( t( t(input) + alpha * other))
}
###########################################
)

```
### subtr_broadcast (for matrix)

```{r}
#' @export
#' 
#'
setGeneric("subtr_broadcast", function(input, other, alpha = 1) {
  standardGeneric("subtr_broadcast")
})
setMethod("subtr_broadcast", "matrix",
###########################################
function(input, other, alpha = 1){
  return(add_broadcast(input, other, (-1*alpha)))
  }
###########################################
)

@@

### curve_normalize
Subtracts the mean function and divides by a scaling constant (deno). This implementation does not use the native fda S3 methods (minus.fd and times.fd), since they do not broadcast well.
```{r}
#' 
#' @export
curve_normalize <- function(functional, mean_functional, deno){
  # operation also takes care of the original_X

@@

### curve_normalize_train_test
Normalizes the functional components of both the training and testing 'predictor_hybrid' objects using the mean and standard deviation derived ONLY from the training data. The scalar components (Z) are left unchanged.

```{r}
#' Normalize Functional Predictors in Hybrid Objects (Train/Test Split)
#' @param train A `predictor_hybrid` object representing the training set.
#' @param test A `predictor_hybrid` object representing the testing set.
#' @return A list containing:
#'   \item{train}{The normalized training `predictor_hybrid` object.}
#'   \item{test}{The normalized testing `predictor_hybrid` object.}
#'   \item{deno_train}{A numeric vector of scaling factors used for each functional predictor.}
#' @export
curve_normalize_train_test <- function(train, test) {
  # --- Type Checks ---
  stopifnot(
    "'train' must be of class 'predictor_hybrid'" = inherits(train, "predictor_hybrid"),
    "'test' must be of class 'predictor_hybrid'" = inherits(test, "predictor_hybrid"),
    "Mismatch in number of functional predictors" = train$n_functional == test$n_functional
  )

@@

```{r}
#'
#' @export
scalar_normalize <- function(scalar_predictor, mean, deno){
   scalar_predictor_normalized <- subtr_broadcast(scalar_predictor, mean) #numerator
   deno_mat <- matrix(rep(1,nrow(scalar_predictor))) %*% deno
   scalar_predictor_normalized <- scalar_predictor_normalized / deno_mat
   return(scalar_predictor_normalized)
 }
```

```{r}
#' Normalize the scalar component (Z) of hybrid predictors using training set statistics.
#'
#' @param predictor_train A `predictor_hybrid` object (train set).
#' @param predictor_test A `predictor_hybrid` object (test set).
#' @return A list containing the normalized 'train' and 'test' hybrid objects.
#' @export
scalar_normalize_train_test <- function(predictor_train, predictor_test){

  # Type and structure checks
  stopifnot(
    inherits(predictor_train, "predictor_hybrid"),
    inherits(predictor_test, "predictor_hybrid"),
    predictor_train$n_scalar == predictor_test$n_scalar
  )

  # Create copies for modification
  predictor_train_normalized <- predictor_train
  predictor_test_normalized <- predictor_test

  # 1. Calculate statistics from the TRAINING set only
  mean_train <- colMeans(predictor_train$Z)
  sd_train <- apply(predictor_train$Z, 2, sd)

  # 2. Apply normalization to the TRAINING set
  predictor_train_normalized$Z <-
    scalar_normalize(
      predictor_train$Z,
      mean_train,
      sd_train
    )

  # 3. Apply normalization to the TEST set using TRAIN statistics
  predictor_test_normalized$Z <-
    scalar_normalize(
      predictor_test$Z,
      mean_train, # IMPORTANT: Use train mean
      sd_train    # IMPORTANT: Use train SD
    )

  return(
    list(
      train = predictor_train_normalized,
      test = predictor_test_normalized
    )
  )
}
```

```{r}
#' --- Dependencies for Testing ---
library(fda) # Required for mock functional data

# Placeholder for a minimal predictor_hybrid constructor for testing purposes

# --- Test Script Start ---

# 1. Setup Mock Data
N_train <- 100
N_test <- 50
N_scalar <- 2
N_functional <- 1

# Create Functional Data (Needs to be present but remain untouched)
basis_obj <- create.bspline.basis(rangeval = c(0, 10), nbasis = 7)
train_coefs <- matrix(rnorm(N_train * 7), nrow = 7)
test_coefs <- matrix(rnorm(N_test * 7), nrow = 7)
mock_fd_train <- fd(train_coefs, basis_obj)
mock_fd_test <- fd(test_coefs, basis_obj)

# Create Scalar Data (Crucial for the test)
# Training data has a clear mean and SD
Z_train <- matrix(rnorm(N_train * N_scalar, mean = c(5, 10), sd = c(2, 3)), 
                  nrow = N_train, ncol = N_scalar, byrow = TRUE)
# Testing data has different properties (must be normalized by train stats)
Z_test <- matrix(rnorm(N_test * N_scalar, mean = c(1, 1), sd = c(10, 10)), 
                 nrow = N_test, ncol = N_scalar, byrow = TRUE)

# Create Hybrid Objects
W_train <- predictor_hybrid(Z = Z_train, functional_list = list(mock_fd_train))
W_test <- predictor_hybrid(Z = Z_test, functional_list = list(mock_fd_test))

# Store original Z for comparison
Z_train_original <- W_train$Z
Z_test_original <- W_test$Z

# 2. Run the Function Under Test
W_normalized_list <- scalar_normalize_train_test(W_train, W_test)
W_train_norm <- W_normalized_list$train
W_test_norm <- W_normalized_list$test

# 3. Verification and Assertions

cat("--- Running Verification Checks ---\n")

# Check 1: Training set is correctly normalized (Mean ~ 0, SD ~ 1)
train_mean_norm <- colMeans(W_train_norm$Z)
train_sd_norm <- apply(W_train_norm$Z, 2, sd)

stopifnot(
  "Train mean Col 1 not near 0" = all(abs(train_mean_norm[1]) < 1e-10),
  "Train mean Col 2 not near 0" = all(abs(train_mean_norm[2]) < 1e-10),
  "Train SD Col 1 not near 1" = all(abs(train_sd_norm[1] - 1) < 1e-10),
  "Train SD Col 2 not near 1" = all(abs(train_sd_norm[2] - 1) < 1e-10)
)
cat("Check 1: Training set Z-score normalization successful (Mean~0, SD~1).\n")

# Check 2: Functional components are untouched
stopifnot(
  "Functional list should not change" = isTRUE(all.equal(
    W_train$functional_list[[1]]$coefs, 
    W_train_norm$functional_list[[1]]$coefs
  ))
)
cat("Check 2: Functional components successfully preserved.\n")

# Check 3: Test set is normalized using TRAIN statistics (crucial consistency check)

# Recalculate the training statistics
train_mean_recalc <- get_mean(Z_train_original)
train_sd_recalc <- get_sd(Z_train_original)

# Manually normalize the test set using TRAIN stats
Z_test_expected <- t( (t(Z_test_original) - train_mean_recalc) / train_sd_recalc)

# Compare the function's output (W_test_norm$Z) against the manual calculation
stopifnot(
  "Test set normalization mismatch" = isTRUE(all.equal(W_test_norm$Z, Z_test_expected))
)
cat("Check 3: Test set correctly normalized using only Training statistics.\n")

cat("--- All tests passed successfully! ---\n")

#
# scalar_normalize_train_test <- function(predictor_train, predictor_test){
#   predictor_train_normalized <- predictor_train
#   predictor_test_normalized <- predictor_test
#
#   mean_train <- get_mean(predictor_train@Z) # train mean function
#   sd_train <- get_sd(predictor_train@Z)
#
#   predictor_train_normalized@Z <-
#     scalar_normalize(
#       predictor_train@Z,
#       mean_train,
#       sd_train
#       )
#
#   predictor_test_normalized@Z <-
#     scalar_normalize(
#       predictor_test@Z,
#       mean_train,
#       sd_train
#     )
#
#   return(
#     list(
#       train = predictor_train_normalized,
#       test = predictor_test_normalized
#     )
#   )
# }
```
